Distributional Trees and Forests

### 2017-12-02 pkg/disttree/inst/CMStatistics2017/slides_rain.Rnw
added references and edited introduction to application (precipitation forecasts)
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@1797 255f4249-2263-4417-b1ea-8498779e27a7

Former-commit-id: 8105fc32dc51df48cc0ba2673eb4610eef9d1460

\title{Distributional Trees and Forests}
\author{Lisa Schlosser, Torsten Hothorn,\\ Heidi Seibold, Reto Stauffer, Achim Zeileis}
\setbeamerfont{url}{size*={11.5pt}{13pt},series=\mdseries}
\URL{https://R-Forge.R-project.org/projects/partykit/}
\headerimage{3}

@@

\begin{frame}[fragile]
\frametitle{Probabilistic Forecasting of Precipitation}
\vspace*{0.2cm}
%Task:
\begin{itemize}
\item Goal: Predict total precipitation amount.
\item Input: Numerical Weather Prediction (NWP).
\item Idea: Statistical Model with numerical predictions as covariates.
\item Data:
\begin{itemize}
\item 95 observation stations in Tyrol.
\item Daily 24h precipitation sums from July of 28 years (1985--2012).
\item Numerical ensemble predictions of meteorological quantities.
\end{itemize}
\item Distribution Assumption: $(\text{Total precipitation})^\frac{1}{1/6} \sim \textit{c}\matchcal{N}$
\end{itemize}
\end{frame}

\begin{frame}[fragile]
\frametitle{Probabilistic Forecasting of Precipitation}
\vspace*{0.2cm}
Variable Selection:
\begin{itemize}
\item Complex terrain \rightarrow unknown influences.
\item 85 Covariates:\\
\begin{itemize}
\item Total precipitation
\item Convective available potential energy
\item Downwards short wave radiation flux (``sunshine'')
\item Mean sea level pressure 
\item Mean sea level pressure difference
\item Preciptable water
\item 2m maximum temperature
\item Total column-integrated condensate
\item Temperature 
\item Temperature differences in altitude
\end{itemize}
\end{itemize}
\end{frame}

@@

\begin{frame}[fragile]
\frametitle{Probabilistic Forecasting of Precipitation}
\vspace*{0.2cm}
Application for all 95 stations:
\begin{itemize}
\item Learn forest model on data from 24 years (1985--2008).
\item Evaluate on 4 years (2009--2012).
\item Compare to commonly used methods:\\
\begin{itemize}
\item Prespecified GAM
\item Boosted GAM
\item Ensemble Model Output Statistics (EMOS) 
\end{itemize}
\end{itemize}
\end{frame}

@@

\begin{frame}
\frametitle{References}

@@

>>>>

\title{Distributional Trees and Forests}
\author{Lisa Schlosser, Torsten Hothorn,\\ Heidi Seibold, Reto Stauffer, Achim Zeileis}
\setbeamerfont{url}{size*={11.5pt}{13pt},series=\mdseries}
\setbeamerfont{footnote}{size=\tiny}
\renewcommand\footnoterule{}
\URL{https://R-Forge.R-project.org/projects/partykit/}
\headerimage{3}

@@

\begin{frame}[fragile]
\frametitle{Probabilistic Forecasting of Precipitation}
\vspace*{-0.1cm}
%Task:
\begin{itemize}
\item Goal: Predict total precipitation amount.
\item Input: Numerical Weather Prediction (NWP).
\item Idea: Statistical model with numerical predictions as \\
\hspace{1cm}covariates.
\item Data:
\begin{itemize}
\item 95 observation stations in Tyrol.
\item Daily 24h precipitation sums from July of 28 years (1985--2012)\footnote{provided by the National Hydrographical Service (EHYD)}.
\item Numerical ensemble predictions of meteorological quantities\footnote{provided by the {U.S.} National Oceanic and Atmospheric Administration (NOAA)\\
\hspace{0.32cm}using the Global Ensemble Forecast System (GEFS)}: 11 members each using slightly different perturbed initial conditions.
\end{itemize}
\item Distribution Assumption:\\
\vspace{0.1cm}
\hspace{1cm}$(\text{Total precipitation})^\frac{1}{1.6} \sim \textit{cens}\mathcal{N}(\mu,\sigma^2)$
\end{itemize}
\end{frame}

\begin{frame}[fragile]
\frametitle{Probabilistic Forecasting of Precipitation}
\vspace*{0.2cm}
Variable Selection:
\begin{itemize}
\item Complex terrain $\rightarrow$ unknown influences.
\item 82 Covariates deduced from 14 base variables \\
(variations such as ensemble minimum/maximum/mean/standard deviation):\\
\begin{itemize}
\item Total precipitation
\item Convective available potential energy
\item Downwards short wave radiation flux (``sunshine'')
\item Mean sea level pressure 
\item Mean sea level pressure difference
\item Preciptable water
\item 2m maximum temperature
\item Total column-integrated condensate
\item Temperature 
\item Temperature differences in altitude
\end{itemize}
\end{itemize}
\end{frame}

@@

\begin{frame}[fragile]
\frametitle{Probabilistic Forecasting of Precipitation}
\vspace*{0.2cm}
Application for all 95 stations:
\begin{itemize}
\item Learn forest model on data from 24 years (1985--2008).
\item Evaluate on 4 years (2009--2012).
\item Compare to commonly used heteroscedastic censored Gaussian models:\\
\begin{itemize}
\item Prespecified GAM: variable selection based on meteorological expert knowledge.
\item Boosted GAM: automatic variable selection.
\item Ensemble Model Output Statistics (EMOS): linear predictors using only total precipitation predictions as covariates.
\end{itemize}
\end{itemize}
\end{frame}

@@

\begin{frame}
\frametitle{References (1)}

@@

\begin{frame}
\frametitle{References (2)}

% Reference for the observation data set
BMLFUW (2016).
  \dquote{{B}undesministerium f\"ur {L}and und {F}orstwirtschaft, {U}mwelt und {W}asserwirtschaft ({BMLFUW}), {A}bteilung {IV/4} - {W}asserhaushalt}
  \text{Available at \url{http://ehyd.gv.at}}, 
  \text{Accessed: {F}ebruary 29 2016}.

% Reference for the GFS reforecast version 2 data set 
Hamill T M, Bates G T, Whitaker J S, Murray D R, Fiorino M, Galarneau T J Jr., Zhu Y, Lapenta W (2013).
  \dquote{{NOAA}'s Second-Generation Global Medium-Range Ensemble Reforecast Dataset}
  \textit{Bulletin of the American Meteorological Society}, 
  \textbf{94}(10), 1553--1565.
  \doi = {10.1175/BAMS-D-12-00014.1}

Messner J W,  Mayr G J, Zeileis A (2016).
  \dquote{Heteroscedastic Censored and Truncated Regression with {crch}}
  \textit{The R Journal}, 
  \textbf{8}(1), 173--181.
  \url{https://journal.R-project.org/archive/2016-1/messner-mayr-zeileis.pdf}

Breiman L (2001).
  \dquote{Random {Forests}}
  \textit{Machine Learning}, 
  \textbf{45}(1), 5--32.
  \doi{10.1023/A:1010933404324}

Hofner B, Mayr A, Fenske N, Schmid M (2017).
  \dquote{{gamboostLSS}: Boosting Methods for {GAMLSS} Models}
  \textit{{R} package version 2.0-0},
  \url{https://CRAN.R-project.org/package=gamboostLSS}

\begin{frame}
\frametitle{References (3)}

\footnotesize
Gneiting T, Raftery A E, Westveld III A H, Goldman T (2005).
  \dquote{Calibrated probabilistic forecasting using ensemble model output statistics and minimum CRPS estimation}
  \textit{Monthly Weather Review}, 
  \textbf{133}(5), 1098--1118.
  \doi{https://doi.org/10.1175/MWR2904.1}

Gebetsberger M, Messner J W, Mayr G J, Zeileis A (2017).
  \dquote{Fine-Tuning Non-Homogeneous Regression for Probabilistic Precipitation Forecasts: Unanimous Predictions, Heavy Tails, and Link Functions}
  \textit{Monthly Weather Review}, 
  \textbf{Forthcoming}.
  \doi{10.1175/MWR-D-16-0388.1}

{CGIAR-CSI} (2016).
  \dquote{{SRTM} 90m Digital Elevation Database {v4.1}}
  \text{Available at \url{http://srtm.csi.cgiar.org}}, 
  \text{Accessed: {J}uly 10 2016}.

Stauffer R, Mayr G J, Messner J W, Umlauf N, Zeileis A (2017).
  \dquote{Spatio-Temporal Precipitation Climatology Over Complex Terrain Using a Censored Additive Regression Model}
  \textit{International Journal of Climatology}, 
  \textbf{37}(7), 3264--3275.
  \doi{10.1002/joc.4913}
